---
title: "NYPD Shooting Data Incident Report"
author: "Isaac Carey"
date: '2022-06-20'
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, tidy.opts = list(width.cutoff = 60), tidy = TRUE)
library(lubridate)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(formatR)
```

## Overview

- Obtain and Preprocess Data
- Examine Missing/Erroneous Data
- Prepare Visualizations & Discuss Bias
- Conclusions and Next Steps

## Missing Values

```{r preprocess, echo = FALSE}

data <- read.csv("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD", na.string = "")
data <- data %>% mutate(Date_Of_Incident = mdy_hms(paste(OCCUR_DATE, OCCUR_TIME, sep = " ")))
data <- data %>% mutate(Hour_Of_Incident = hour(Date_Of_Incident))       
na_counts <- data %>% select(everything()) %>% summarise_all(funs(sum(is.na(.))))
```
Missing values seem reasonable in this context.
```{r NAs, echo = FALSE}
colSums(na_counts)
```

## Erroneous Values

```{r typos, echo = FALSE}
table(data$PERP_AGE_GROUP)
```
- Impute? Remove? 

- Any introduction of bias?

## First Consideration: Time of Incident

- Frequency chart by hour will give an easy snapshot of the times of incidents.

- Likely a predictable result, but can inform/reinforce scheduling of officers.

## Visualization of Incident Times

```{r incidentgraph, echo = FALSE}
ggplot(data, aes(x= Hour_Of_Incident)) + geom_bar() + labs(x = "Hour of Incident", y = "Frequency")
```


## Second Consideration: Age of Perpetrators

- Erroneous values removed with no major bias concerns.

- High values of a particular age group can lead to more targeted outreach.

- Unknown vs NA

## Graph of Age Distribution of Perpetrators

```{r agegraph, echo = FALSE}
filtered_ages <- data %>% filter(PERP_AGE_GROUP %in% c("<18", "18-24", "25-44", "45-64", "65+", "UNKNOWN"))
ggplot(filtered_ages, aes(x= factor(PERP_AGE_GROUP))) + geom_bar() + labs(x = "Perp Age Group", y = "Frequency")
```

- Potential bias issue.

## Third Consideration: Breakdown of Incidents by Borough

- This can also help the city to better allocate its resources.

- Is this a biased view?


## Table of Incidents by Borough

```{r boroughgraph, echo = FALSE}
ggplot(data, aes(x = BORO)) + geom_bar() + labs(x = "Borough", y = "Frequency")
```

## Conclusions and Next Steps

- Potentially consider relative incident rates by borough instead of total amounts.

- Consider victim/perpetrator age difference.

- Pilot different schedules of duty for officers to address nighttime incidents.

## Thanks!

Thank you so much for taking the time to go through this.
